Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-15 paper/sheaf_lics.tex
first complete pass on section 3

As noticed in \cite{hottbook}, left exact modalities corresponds 
to sub-$(\infty,1)$-topoi.

A left exact modality leads to a new homotopy type theory with
potentially new principles.

Definition of Sigma-Types.

Preservation of univalence.

\nt{state the property on IsEquiv to show preservation of univalence}

Every left exact modality induces a new type theory, but with more
properties, that can make it inconsistent. We now provide a simple
characterization of left exact modalities that give rise to a
consistent theory.

@@

A basic example is the {\em open modality of $P$} for a $P:\HProp$,
$\modal_P T = P \to T$.  When one works in the subuniverse defined by
this modality, defining an inhabitant of a type $T$ is actually giving
an inhabitant of $P \to T$. Thus, being in this subuniverse is just
adding the axiom $P$ to the theory. Sadly, adding an axiom this way
does not add any computational content to $P$, thus is not completely
satisfactory.

Another example is the $\lnot\lnot$-modality, $\modal_{\lnot\lnot} T = (T
\to \bot) \to \bot$. Although we might think it enables us to use
classical reasonings in the corresponding reflective subuniverse, the
main problem is that every type in the new universe is an $\HProp$,
which is very restrictive. 
We will use this modality only in the $\HProp$ layer, and try to
extend it to every $\Type_n$ without collapsing every type.

>>>>

As noticed in \cite{hottbook}, left exact modalities corresponds 
to sub-$(\infty,1)$-topoi. Indeed, modal types defines a new type
theory as modality preserves all type former, such as dependent
products, dependent sums or inductive types (\nt{double check that}).

Furthermore, when $A$ and $B$ are modal and $f : A \to B$, the mere
proposition of $f$ being an equivalence is also modal. This means that
univalence also holds in $\Type^\modal_p$, so it induces a good notion
of homotopy type theory.

Every left exact modality induces a new homotopy type theory, but with more
properties, that can make it inconsistent. We now provide a simple
characterization of left exact modalities that give rise to a
consistent theory.

@@

A first example of left exact modality for every level $p$ is the {\em
  open modality of $P$} for a $P:\HProp$ defined as
$\modal_P T = P \to T$.
%
When one works in
the subuniverse defined by this modality, defining an inhabitant of a
type $T$ amounts to give an inhabitant of $P \to T$. 
%
Thus, being in this subuniverse corresponds to adding the axiom $P$ to the
theory. Of course, adding an axiom this way does not add any computational
content to $P$, and the consistency of the modallity corresponds to
the fact that $P \to \zero$ (or $\lnot P$) is not inhabited.

Another example of left exact modality for every level $p$ is the
double negation modality,
$\modal_{\lnot\lnot} T = (T \to \zero) \to \zero$.
%
This modality enables to use
classical reasonings in the corresponding reflective subuniverse, but
every type in the new universe is collapsed to an $\HProp$, ending up
with porpositional classical logic.

In this paper, we use the double negation modality at level $-1$ only,
and then extend it to higher homotopy types using sheafification. This
way, we can add classical reasoning in the type theory without
collapsing the theory to mere propositions.
